Constraints on the quasi-normal mode frequencies of the LIGO-Virgo signals by making full use of gravitational-wave modeling

### 2021-02-02 papers/intro_paper/results_lvc.tex
addressed remaining comments in results sections

The LIGO-Virgo Collaboration recently released their testing GR
catalogue containing results for events observed
during O3a~\cite{Abbott:2020jks}. For the test that we here present, the results shown in~\cite{Abbott:2020jks} only include events which pass a threshold for the median redshifted total mass $\geq 90 \Mo$ and SNRs in the pre- and
post-merger regions $\geq 8$\footnote{The pre- and post-merger regions of the signal are identified from the signal's power before and after the signal reaches the peak's amplitude, which is determined from the maximum of the likelihood function obtained with the parameter-estimation analysis.}. 
%
The SNR threshold ensures that the signal contains sufficient information in both the inspiral and merger stages to break the degeneracy between the binary's total mass and the non-GR deviations $(\df{220}, \dtau{220})$. Such strong degeneracy is present for low-SNR events with negligible higher-modes and for which only the post-merger is detectable rendering the measurement of $(\df{220}, \dtau{220})$ impossible for those cases.\rb{should we show an example where this degeneracy happens in the appendix? We can show a corner plot with $(\df{220}, \dtau{220}, M_{\rm total })$ for 21g in the appendix for example. And perhaps also for GW150914 for comparison.}
On the other hand the total mass threshold $\geq 90 \Mo$ was employed due to the fact that this analysis is computationally expensive, and also because we expected these events to be the most promising for ringdown studies. However, since the SNR threshold alone should be sufficient for the analysis, for this paper we run the test on all the events listed in Ref.~\cite{Abbott:2020jks} that have SNRs in the pre- and post-merger regions $\geq 8$, without imposing any mass threshold.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% O1-O2 events
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure*}
        \includegraphics[width=0.5\textwidth]{figures/rin_pseob_results_v2.pdf}\includegraphics[width=0.3\textwidth]{figures/rin_all_events_bounds.pdf}
        \caption{\emph{Top panel}:\rb{should this be left and right panel?} The 90\% credible levels of the posterior probability distribution of the fractional deviations in the frequency and damping time of the $(2,\pm 2)$ QNM, $(\df{220},\dtau{220})$ and their corresponding one-dimensional marginalized posterior distributions, for events from O1, O2 and O3a passing a SNR threshold of $8$ in both the pre- and post-merger signal. The solid red curve marks the best single-event constraint, GW150914, whereas the contraints from the other events are indicated by the dash-dot curves. The joint constraints on $(\df{220},\dtau{220})$ obtained multiplying the likelihoods from individual events is given by the filled grey contours, while the hierarchical method of combination yields the black dot dashed curves. \emph{Bottom panel}: 90\% credible interval on the one-dimensional marginalised posteriors on $\delta \sigma_i=(\df{220},\dtau{220})$, colored by the median redshifted total mass $(1 + z)M$, inferred assuming GR. Filled gray (unfilled black) triangles mark the constraints obtained when all the events are combined by multiplying likelihoods (hierarchically).}
        \label{fig:o1o2_events}
\end{figure*}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

For all the relevant signals, we show the posterior distributions $(\df{220}, \dtau{220})$ in the left panel \rb{the caption says "top panel" but the current figure is in the left panel} of Fig.~\ref{fig:o1o2_events} and also provide the reconstructed QNM parameters, $(\fngr{220}, \taungr{220})$ in Table~\ref{tab:qnm_o1o2_results}. In the right panel \rb{the caption says "bottom panel" but the current figure is in the right panel} of Fig.~\ref{fig:o1o2_events} we also provide a summary of the 90\% credible intervals on the one-dimensional marginalised posteriors highlighting the dependence of the constraints with respect to the total mass of the system. In general the tightest bounds come from the most massive systems, as would be expected given that more massive events tend to have larger post-merger SNR \rb{check if this is true when plot is updated}.

Among all the GW signals detected so far, GW150914 (solid curve in Fig.~\ref{fig:o1o2_events}) is unique in its loudness, leading to the first, and to date, best attempt in measuring the QNM frequencies~\cite{LSC_2016grtests,Brito:2018rfr,Carullo:2019flw,Isi:2019aib}. Within 90\% credibility we obtain from GW150914: 
%
\begin{equation}
\df{220}=0.04^{+0.06}_{-0.04}\,\quad \dtau{220}=0.09^{+0.18}_{-0.18}\,.
\end{equation}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Table for LVC values
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{table}
\input{table_lvc_results}
\caption{\textcolor{red}{NOT COMPLETE} \comment{AB: We would need to give also the final mass and spins. I would suggest that we list in the Table
also the results from the TGR GWTC-2 paper, obtained with our method, so that all the results are in one Table.}}
\label{tab:qnm_o1o2_results}
\end{table}

>>>>

The LIGO-Virgo Collaboration recently released their testing GR
catalogue containing results for events observed
during O3a~\cite{Abbott:2020jks}. For the test that we here present, the results shown in~\cite{Abbott:2020jks} only include events which pass a threshold for the median redshifted total mass $\geq 90 \Mo$ and SNRs in the pre- and
post-merger regions $\geq 8$\footnote{The pre- and post-merger regions of the signal are identified from the signal's power before and after the signal reaches the peak's amplitude, which is determined from the maximum of the likelihood function obtained with the parameter-estimation analysis.}. 
%
The SNR threshold ensures that the signal contains sufficient information in both the inspiral and merger stages to break the degeneracy between the binary's total mass and the non-GR deviations $(\df{220}, \dtau{220})$. Such strong degeneracy is present for low-SNR events with negligible higher-modes and for which only the post-merger is detectable rendering the measurement of $(\df{220}, \dtau{220})$ impossible for those cases.\rb{should we show an example where this degeneracy happens in the appendix? We can show a corner plot with $(\df{220}, \dtau{220}, M_{\rm total })$ for 21g in the appendix for example. And perhaps also for GW150914 for comparison.}\abhi{I would be fine with this. Perhaps we should have a discussion about how best to put it in the appendix, if we decide to.}
On the other hand the total mass threshold $\geq 90 \Mo$ was employed due to the fact that this analysis is computationally expensive, and also because we expected these events to be the most promising for ringdown studies. However, since the SNR threshold alone should be sufficient for the analysis, for this paper we run the test on all the events listed in Ref.~\cite{Abbott:2020jks} that have SNRs in the pre- and post-merger regions $\geq 8$, without imposing any mass threshold.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% O1-O2 events
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure*}
        \includegraphics[width=0.5\textwidth]{figures/rin_pseob_results_v2.pdf}\includegraphics[width=0.3\textwidth]{figures/rin_all_events_bounds.pdf}
        \caption{\emph{Left panel}:The 90\% credible levels of the posterior probability distribution of the fractional deviations in the frequency and damping time of the $(2,\pm 2)$ QNM, $(\df{220},\dtau{220})$ and their corresponding one-dimensional marginalized posterior distributions, for events from O1, O2 and O3a passing a SNR threshold of $8$ in both the pre- and post-merger signal. The solid red curve marks the best single-event constraint, GW150914, whereas the contraints from the other events are indicated by the dash-dot curves. The joint constraints on $(\df{220},\dtau{220})$ obtained multiplying the likelihoods from individual events is given by the filled grey contours, while the hierarchical method of combination yields the black dot dashed curves. \emph{Right panel}: 90\% credible interval on the one-dimensional marginalised posteriors on $\delta \sigma_i=(\df{220},\dtau{220})$, colored by the median redshifted total mass $(1 + z)M$, inferred assuming GR. Filled gray (unfilled black) triangles mark the constraints obtained when all the events are combined by multiplying likelihoods (hierarchically).}
        \label{fig:o1o2_events}
\end{figure*}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

For all the relevant signals, we show the posterior distributions $(\df{220}, \dtau{220})$ in the left panel of Fig.~\ref{fig:o1o2_events} and also provide the reconstructed QNM parameters, $(\fngr{220}, \taungr{220})$ in Table~\ref{tab:qnm_o1o2_results}. In the right panel  of Fig.~\ref{fig:o1o2_events} we also provide a summary of the 90\% credible intervals on the one-dimensional marginalised posteriors highlighting the dependence of the constraints with respect to the total mass of the system. In general the tightest bounds are expected to come from the most massive systems, as they tend to have larger post-merger SNR. We find a similar trend in the right panel of Fig.~\ref{fig:o1o2_events}.

Among all the GW signals detected so far, GW150914 (solid curve in Fig.~\ref{fig:o1o2_events}) is unique in its loudness, leading to the first, and to date, best attempt in measuring the QNM frequencies~\cite{TheLIGOScientific:2016src,Brito:2018rfr,Carullo:2019flw,Isi:2019aib}. Within 90\% credibility we obtain from GW150914: 
%
\begin{equation}
\df{220}=0.04^{+0.06}_{-0.04}\,\quad \dtau{220}=0.09^{+0.18}_{-0.18}\,.
\end{equation}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Table for LVC values
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{table}
\input{table_lvc_results}
\caption{The median, and symmetric 90\% credible intervals of the frequency and damping time of the $(2,\pm2)$ QNM, and the mass and spin of the remnant object estimated from the measurements of the QNM frequencies.}
\label{tab:qnm_o1o2_results}
\end{table}
